### 2017-05-18 sheets/basis-2/sheet-basis-2.tex
Add solutions for exercises by Dominik

\DeclareMathOperator{\A}{\mathscr A}
\DeclareMathOperator{\B}{\mathscr B}
\DeclareMathOperator{\M}{\mathscr M}

@@

\begin{answer}
  \begin{enumerate}
  \item A subcategory of an ordered set $(S,\leq)$ is an ordered set $(S',\leq')$ such that $S'\subseteq S$ and $\leq'\;\subseteq\;\leq$. In the case of full subcategories even $\leq'\;=\;\leq$ holds.
  \item The subcategories of a group are either the category with no objects or submonoids of the group (which is of course a monoid). The only full subcategories are the group itself and the category with no objects. 
  \end{enumerate}
\end{answer}

@@

@@

@@

@@

>>>>

\newcommand{\A}{\cat A}
\newcommand{\B}{\cat B}
\newcommand{\M}{\cat M}
\newcommand{\Rel}{\fcat{Rel}}

@@

\begin{answer}
  \begin{enumerate}
  \item A subcategory of an ordered set $(S,\leq)$ is an ordered set $(S',\leq')$ such that $S'\subseteq S$ and $\leq'\;\subseteq\;\leq$. In the case of full subcategories even for all $s,t\in S'$, $s\leq t$ if and only if $s\leq' t$ holds.
  \item The subcategories of a group are either the category with no objects or submonoids of the group (which is of course a monoid). The only full subcategories are the group itself and the category with no objects.
  \end{enumerate}
\end{answer}

@@

\begin{answer}
  For every arrow $x \to y$ in $\A$, there is an arrow $f : y \to x$ in $\A^\op$ and hence an arrow $F(f) : F(y) \to F(x)$ in $\B$.
  In the language of posets this means that $x \leq y$ implies $f(x) \geq f(y)$, i.e.\ $f$ is monotonically decreasing.
  The functor laws hold tautologically in posets and don't tell us anything about $f$.
\end{answer}

@@

\begin{answer}
  \begin{enumerate}
    \item The categories are isomorphic.
      We define functors $F : \Rel^\op \to \Rel$ and $G : \Rel \to \Rel^\op$ in both directions by
      \begin{align*}
        F(X) &\coloneqq X  &  G(X) &\coloneqq X \\
        F(R) &\coloneqq R^{-1}  &  G(R) &\coloneqq R^{-1}
      \end{align*}
      for all sets $X$ and $Y$ and relations $R \in X \times Y$.
      Both functors are mutually inverse bacause $(R^{-1})^{-1} = R$.
      Thus $\Rel^\op \iso \Rel$.

    \item We refute the statement by contradiction.
      Assume that we have an isomorphism $F : \Set^\op \to \Set$.
      Let $1$ be a one-element set.
      It is final in $\Set$ and hence initial in $\Set^\op$ and will be mapped to the initial object $\emptyset$ in $\Set$, because isomorphisms preserve initial objects.
      We define $f : 1 \to X$ as a function to some set $X$ that is not isomorphic to $1$.
      Then $F(X)$ is not isomorphic to $F(1) = \emptyset$, but there is a function $F(f) : F(X) \to \emptyset$, which leads to a contradiction.

    \item Let $X$ be a finite set with at least two elements.
      We define the monoid $\M \coloneqq \lbrace f : X \to X \rbrace$ with function composition as binary operation and the identity function as neutral element.
      Let $c \in \M$ be some constant function.
      Such a function is left-absorbing, which means that for all functions $f \in \M$, we have $c \circ f = c$.
      Left-absorbing elements in the opposite monoid correspond to right-absorbing functions, i.e., fuctions $d \in \M$ such that for all $f \in M$, we have $f \circ d = d$.
      Obviously such functions do not exist.
      Since left-absorbing elements are preserved by isomorphisms, $\M$ cannot be isomorphic to its opposite monoid.

    \item We define mutually inverse functors $F : \CAT_i^\op \to \CAT_i$ and $G : \CAT_i \to \CAT_i^\op$ by
      \begin{align*}
        F(\A) &\coloneqq \A  &  G(\A) &\coloneqq \A \\
        F(H) &\coloneqq H^{-1}  &  G(H) &\coloneqq H^{-1}
      \end{align*}
      for all categories $\A$ and $\B$ and functors $H : \A \to B$.
      The functors $F$ and $G$ are mutually inverse because $(H^{-1})^{-1} = H$.
      Thus $\CAT_i^\op \iso \CAT_i$.
  \end{enumerate}
\end{answer}

@@

\begin{answer}
  We define a functor $F : \Set_* \to 1/\Set$ by
  \begin{align*}
    F((A, a)) &\coloneqq (x \mapsto a) \\
    F(h) &\coloneqq h
  \end{align*}
  for all pointed sets $(A, a)$ and $(B, b)$ and functions $h : A \to B$ with $f(a) = b$.
  The second line is valid because
  \[ \begin{split}
    h \circ F((A, a))
    &= (x \mapsto h(a)) \\
    &= (x \mapsto b) \\
    &= F((B, b)).
  \end{split} \]
  Now we define the inverse functor $G : 1/\Set \to \Set_*$ by
  \begin{align*}
    G(f) &\coloneqq (A, f(\star)) \\
    G(h) &\coloneqq h
  \end{align*}
  for all functions $f : 1 \to A$, $g : 1 \to B$ and $h : A \to B$ with $h \circ f = g$.
  The second line is valid because
  \[ \begin{split}
    h(f(\star))
    &= (h \circ f)(\star) \\
    &= g(\star).
  \end{split} \]
  Both functors are mutually inverse on objects because
  \[ \begin{split}
    F(G(f))
    &= F((A, f(\star))) \\
    &= (x \mapsto f(\star)) \\
    &= f
  \end{split} \]
  and
  \[ \begin{split}
    G(F((A, a)))
    &= G(x \mapsto a) \\
    &= (A, a).
  \end{split} \]
  On arrows they are trivially inverse.
\end{answer}

@@

\begin{answer}
  As suggested, we define
  \begin{align*}
    F(A) &\coloneqq \A / A \\
    F(f) &\coloneqq F_f,
  \end{align*}
  where $F_f : \A / A \to \A / B$ is given by
  \begin{align*}
    F_f(g) &\coloneqq f \circ g \\
    F_f(h) &\coloneqq h
  \end{align*}
  for all functions $g$, $g'$ and $h$ such that the following diagram commutes.
  \[ \begin{tikzcd}[column sep=tiny]
    C \arrow[dr, "g"'] && C' \arrow[ll, "h"] \arrow[dl, "g'"] \\
    & A \arrow[d, "f"] \\
    & B
  \end{tikzcd} \]
  The second line is valid because
  \[ \begin{split}
    F_f(g) \circ h
    &= f \circ g \circ h \\
    &= f \circ g' \\
    &= F_f(g')
  \end{split} \]
  Our definition of $F$ satisfies both functor laws:
  \begin{enumerate}
    \item For $f : A_2 \to A_3$ and $f' : A_1 \to A_2$, we have $F(f \circ f') = F(f) \circ F(f')$:
      \begin{align*}
        F_{f \circ f'}(g)
          &= f \circ f' \circ g \\
          &= F_f(F_f'(g)) \\
        F_{f \circ f'}(h)
          &= h \\
          &= F_f(F_f'(h))
      \end{align*}
    \item We have $F(1_A) = 1_{F(A)}$:
      \begin{align*}
        F_{1_A}(g) &= g \\
        F_{1_A}(h) &= h
      \end{align*}
  \end{enumerate}
  In general the functor $F$ is not an isomorphism.
  For example, if $\A$ is the empty category then it is not isomorphic to $\Cat$.
\end{answer}
